### 2019-09-20 2019_neurips_kr2ml/summary.tex

\section{Conclusions}

In summary, the results clearly indicate that the mechanisms introduced in SAMNet  enable it to learn to operate independently of the total number of frames or number of distractions, and allow it to generalize to longer videos and more complex scenes. One other strength of SAMNet is its interpretability. Observing attention maps (see Appendix) shows that SAMNet can effectively perform multi-step reasoning over questions and frames as intended. It also accurately classifies temporal contexts as designed. However we notice that the model can sometime discover alternative strategies that were not in the intended design but the answers are still correct. More details are given in the Appendix.

Summarizing, the system has learned to reason using neural representations similarly as a human would operate on abstract symbols when solving the same task, including objects that are memorized and recalled from the memory.