### 2019-11-15 cvpr_submission/paper_main.tex
Removed unnecessary packages + slightly reformulated abstract. Abstract probably needs a little more work.

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage[font=small]{subcaption}
\captionsetup[figure]{skip=5pt}
\captionsetup[table]{skip=5pt}
\setlength{\belowcaptionskip}{-15pt}  % reduce padding below figure
\usepackage{amsmath, amssymb, amsfonts,bm}
\usepackage{amsthm}

% Include other packages here, before hyperref.
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{tabularx}

\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

%\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{7813} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi

\usepackage{cleveref}
\crefname{section}{Section}{Sections}
\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{appendix}{Appendix}{Appendices}
\crefname{definition}{Definition}{Definitions}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem*{notation}{Notation}
%\newtheorem*{remark}{Remark}
\newtheorem*{note}{Note}

\newtheorem{fact}{Fact}
\newtheorem*{observation}{Observation}
\newtheorem*{condition}{Condition}
\newtheorem*{claim}{Claim}
\newtheorem*{example}{Example}
\newtheorem*{question}{Question}

\usepackage{enumitem}
\setlist{leftmargin=*}

%% User's macros %%
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\cF}{\mathcal{F}}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vc}{\vec{c}}
\newcommand{\vso}{\vec{so}}
\newcommand{\vone}{\vec{1}}
\newcommand{\proj}{\mathrm{proj}}
\newcommand{\vx}{\vec{x}}
\newcommand{\vy}{\vec{y}}

\newcommand{\imatch}{g^{\mathrm{v}}}
\newcommand{\mmatch}{g^{\mathrm{m}}}

\newcommand{\doadd}{h^{\mathrm{a}}}
\newcommand{\doreplace}{h^{\mathrm{r}}}

\newcommand{\tlast}{\tau^{\mathrm{last}}}
\newcommand{\tlatest}{\tau^{\mathrm{latest}}}
\newcommand{\tnow}{\tau^{\mathrm{now}}}
\newcommand{\tnone}{\tau^{\mathrm{none}}}

\newcommand{\rhead}{\vec{rh}}
\newcommand{\whead}{\vec{wh}}

\newcommand{\tk}[1]{\textcolor{red}{TK: #1}}
\newcommand{\ao}[1]{\textcolor{green}{AO: #1}}
\newcommand{\tsj}[1]{\textcolor{magenta}{TSJ: #1}}
\newcommand{\vm}[1]{\textcolor{blue}{VM: #1}}

\newcommand{\compresslist}{ % Define a command to reduce spacing within itemize/enumerate environments, this is used right after \begin{enumerate} or \begin{enumerate}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}}

%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%% TITLE
\title{Transfer Learning in Multimodal Compositional and Relational Reasoning}

%%%%%%%%% AUTHORS
% WILL NEED TO REARRANGE THE AUTHORS LIST & GET CORRECT INSTITUTIONS / EMAILS
\author{T.S. Jayram \and  Vincent Marois \and Vincent Albouy \and Tomasz Kornuta  \and Emre Sevgen \and Ahmet Ozcan\\
IBM Research AI\\
Almaden Research Center, San Jose, CA 95120, USA\\
Thomas J. Watson Research Center, Yorktown Heights, NY 10598, USA\\
{\tt\small {jayram,  asozcan}@us.ibm.com}
{\tt\small {tkornuta, sesevgen}@gmail.com}
{\tt\small vincent.marois@ibm.com}
}

\maketitle

%%%%%%%%% ABSTRACT
\begin{abstract}
  Transfer learning is becoming the de facto solution for vision and text encoders in the front-end processing of machine learning solutions. Utilizing vast amounts of knowledge in pre-trained models and subsequent fine-tuning allows achieving better performance in domains where labeled data is limited. In this paper, we analyze the efficiency of transfer learning in multimodal (vision and language) reasoning by introducing a new model (SAMNet) and testing it on two datasets: COG and CLEVR.  Our new model achieves state-of-the-art accuracy on COG and shows significantly better generalization capabilities compared to the baseline. We also formalize a taxonomy of transfer learning for visual reasoning around three axes: feature, temporal, and reasoning transfer. Based on extensive experimentation on transfer learning on the two datasets, we show the performance of the new model along each axis.
\end{abstract}

%%%%%%%%% BODY TEXT
\input{introduction}

\input{related_work}

\input{transfer_learning}

\input{samnet}

\input{experiments}

\input{summary}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{bibliography}
}

\clearpage
\input{appendix}

\end{document}