### 2019-11-15 cvpr_submission/supplementary_material.tex
Made Supplementary Material its own file.

\documentclass[10pt,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{url}
\usepackage{enumitem}
\setlist{leftmargin=*}

%\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{7813} 
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\ifcvprfinal\pagestyle{empty}\fi

\begin{document}

%%%%%%%%% TITLE
\title{Transfer Learning in Multimodal Compositional and Relational Reasoning:\\Supplementary Material}

%%%%%%%%% AUTHORS
% WILL NEED TO REARRANGE THE AUTHORS LIST & GET CORRECT INSTITUTIONS / EMAILS
\author{T.S. Jayram \and  Vincent Marois \and Vincent Albouy \and Tomasz Kornuta  \and Emre Sevgen \and Ahmet Ozcan\\
IBM Research AI\\
Almaden Research Center, San Jose, CA 95120, USA\\
Thomas J. Watson Research Center, Yorktown Heights, NY 10598, USA\\
{\tt\small {jayram,  asozcan}@us.ibm.com}
{\tt\small {tkornuta, sesevgen}@gmail.com}
{\tt\small vincent.marois@ibm.com}
}

\maketitle

\input{appendix}

\clearpage
{\small
	\bibliographystyle{ieee_fullname}
	\bibliography{bibliography}
}

\end{document}