### 2019-11-20 Transfer_Learning/arxiv_version/paper_main.tex
Rearranging folders because of arxiv version

\documentclass[11pt]{article}

\usepackage[letterpaper,margin=1in]{geometry}
%\usepackage{cvpr}
%\usepackage{times}
%\usepackage{epsfig}
\usepackage{graphicx}
\usepackage[font=small]{subcaption}
\captionsetup[figure]{skip=5pt}
\captionsetup[table]{skip=5pt}
\setlength{\belowcaptionskip}{-15pt}  % reduce padding below figure
\usepackage{amsmath, amssymb, amsfonts,bm}
\usepackage{amsthm}

% Include other packages here, before hyperref.
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{tabularx}
\usepackage{xcolor}

\usepackage{hyperref}
\hypersetup{
	pagebackref,
	colorlinks,
	breaklinks,
	bookmarks=false,
	linkcolor={red!50!black},
	citecolor={blue!50!black},
	urlcolor={blue!80!black},
	%linkcolor=OliveGreen,
	%citecolor=Blue,
}

\usepackage{cleveref}
\crefname{section}{Section}{Sections}
\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{appendix}{Appendix}{Appendices}
\crefname{definition}{Definition}{Definitions}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}

\theoremstyle{remark}
\newtheorem*{notation}{Notation}
%\newtheorem*{remark}{Remark}
\newtheorem*{note}{Note}

\newtheorem{fact}{Fact}
\newtheorem*{observation}{Observation}
\newtheorem*{condition}{Condition}
\newtheorem*{claim}{Claim}
\newtheorem*{example}{Example}
\newtheorem*{question}{Question}

\usepackage{enumitem}
\setlist{leftmargin=*}

%% User's macros %%
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\cF}{\mathcal{F}}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vc}{\vec{c}}
\newcommand{\vso}{\vec{so}}
\newcommand{\vone}{\vec{1}}
\newcommand{\proj}{\mathrm{proj}}
\newcommand{\vx}{\vec{x}}
\newcommand{\vy}{\vec{y}}

\newcommand{\imatch}{g^{\mathrm{v}}}
\newcommand{\mmatch}{g^{\mathrm{m}}}

\newcommand{\doadd}{h^{\mathrm{a}}}
\newcommand{\doreplace}{h^{\mathrm{r}}}

\newcommand{\tlast}{\tau^{\mathrm{last}}}
\newcommand{\tlatest}{\tau^{\mathrm{latest}}}
\newcommand{\tnow}{\tau^{\mathrm{now}}}
\newcommand{\tnone}{\tau^{\mathrm{none}}}

\newcommand{\rhead}{\vec{rh}}
\newcommand{\whead}{\vec{wh}}

\newcommand{\uX}{\underline{X}}

\newcommand{\tk}[1]{\textcolor{red}{TK: #1}}
\newcommand{\ao}[1]{\textcolor{green}{AO: #1}}
\newcommand{\tsj}[1]{\textcolor{magenta}{TSJ: #1}}
\newcommand{\vm}[1]{\textcolor{blue}{VM: #1}}

\newcommand{\compresslist}{ % Define a command to reduce spacing within itemize/enumerate environments, this is used right after \begin{enumerate} or \begin{enumerate}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}}

%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%% TITLE
\title{Transfer Learning in Visual and Relational Reasoning}

%%%%%%%%% AUTHORS
% WILL NEED TO REARRANGE THE AUTHORS LIST & GET CORRECT INSTITUTIONS / EMAILS
\author{T.S. Jayram \and  Vincent Marois \and \and Tomasz Kornuta \and Vincent Albouy \and Emre Sevgen \and Ahmet Ozcan\\
IBM Research AI\\
Almaden Research Center, San Jose, CA 95120, USA\\
Thomas J. Watson Research Center, Yorktown Heights, NY 10598, USA\\
{\tt\small \{jayram,  asozcan\}@us.ibm.com}
{\tt\small \{tkornuta, sesevgen\}@gmail.com}
{\tt\small vincent.marois@ibm.com}
}

\maketitle

%%%%%%%%% ABSTRACT
\begin{abstract}
  Transfer learning is becoming the de facto solution for vision and text encoders in the front-end processing of machine learning solutions. Utilizing vast amounts of knowledge in pre-trained models and subsequent fine-tuning allows achieving better performance in domains where labeled data is limited. In this paper, we analyze the efficiency of transfer learning in visual reasoning by introducing a new model (SAMNet) and testing it on two datasets: COG and CLEVR.  Our new model achieves state-of-the-art accuracy on COG and shows significantly better generalization capabilities compared to the baseline. We also formalize a taxonomy of transfer learning for visual reasoning around three axes: feature, temporal, and reasoning transfer. Based on extensive experimentation of transfer learning on each of the two datasets, we show the performance of the new model along each axis.
\end{abstract}

%%%%%%%%% BODY TEXT
\input{introduction}

\input{related_work}

\input{transfer_learning}

\input{samnet}

\input{experiments}

\input{summary}

\bibliographystyle{alpha}
\bibliography{bibliography}

\newpage
\input{appendix}

\end{document}