### 2019-11-21 Transfer_Learning/cvpr_submission/supplementary_material.tex

\documentclass[10pt,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}

\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{xcolor}

\usepackage{hyperref}
\hypersetup{
	colorlinks,
	breaklinks,
	linkcolor={red!50!black},
	citecolor={blue!50!black},
	urlcolor={blue!80!black},
}

\usepackage{cleveref}
\crefname{table}{Table}{Tables}

%\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{7813} 
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\ifcvprfinal\pagestyle{empty}\fi

\begin{document}

%%%%%%%%% TITLE
\title{Transfer Learning in Visual and Relational Reasoning\\ (Supplementary Material)}

%%%%%%%%% AUTHORS
\author{T.S. Jayram \and  Vincent Marois \and Vincent Albouy \and Tomasz Kornuta  \and Emre Sevgen \and Ahmet Ozcan\\
IBM Research AI\\
Almaden Research Center, San Jose, CA 95120, USA\\
Thomas J. Watson Research Center, Yorktown Heights, NY 10598, USA\\
{\tt\small {jayram,  asozcan}@us.ibm.com}
{\tt\small {tkornuta, sesevgen}@gmail.com}
{\tt\small vincent.marois@ibm.com}
}

\maketitle

\input{appendix}

\bigskip

\clearpage
\bibliographystyle{ieee_fullname}
\bibliography{bibliography}

\end{document}