### 2019-11-22 Transfer_Learning/cvpr_submission/appendix.tex
Fixed appendix in both supplementary and arxiv. Added results from SoftPaths to the arxiv version. Added a comment in the main body as well.

\section{Datasets: CLEVR-CoGenT and COG}
\label{sec:datasets-desc}
%Here are the details of the COG dataset.
%
The CoGenT dataset~\cite{johnson2017clevr} contains:
\begin{itemize}
	\def\labelitemi{--}
	\item Training set of 70,000 images and 699,960 questions in Condition A,
	\item Validation set of 15,000 images and 149,991 questions in Condition A,
	\item Test set of 15,000 images and 149,980 questions in Condition A (without answers),
	\item Validation set of 15,000 images and 150,000 questions in Condition B,
	\item Test set of 15,000 images and 149,992 questions in Condition B (without answers),
	\item Scene graphs and functional programs for all training/validation images/questions.
\end{itemize}

The combinations of attributes in CoGenT-A and CoGenT-B are shown in~\cref{tab:cogent_conditions_supplement}.
\begin{table}[ht]
	\centering
	\begin{tabular}{cccc}
		\toprule
		Dataset	&	Cubes	&	Cylinders	&	Spheres	\\
		\midrule
		CoGenT-A	&	Gray / Blue / Brown / Yellow	&	Red / Green / Purple / Cyan	&	Any color	\\
		CoGenT-B	&	Red / Green / Purple / Cyan	&	Gray / Blue / Brown / Yellow	&	Any color 	\\
		\bottomrule
	\end{tabular}
	\caption{Colors \& shapes combinations present in CoGenT-A \& -B datasets.}
	\label{tab:cogent_conditions_supplement}
\end{table}

The Canonical and Hard variants of the COG dataset~\cite{yang2018dataset} are contrasted in~\cref{tab:cog_variants_supplement}.
\begin{table}[ht]
	\centering
	\begin{tabular}{lcccccc}
		\toprule
		Variant    &  	number &  	maximum & maximum & size & size & size  \\ 
		& of   & memory & number of & of & of & of  \\
		& frames & duration & distractors & training set & validation set & test set \\
		\midrule
		Canonical & 4 & 3 & 1 & 10000320 & 500016 & 500016 \\	
		Hard  & 8 & 7 & 10 & 10000320 & 500016  & 500016 \\
		\bottomrule	
	\end{tabular}
	\caption{Details of the Canonical and Hard variants of the COG dataset.}
	\label{tab:cog_variants_supplement}
\end{table}
%
%\begin{table}[ht]
%	\centering
%	\begin{tabular}{lcccccc}.
%		\toprule
%		Variant    &  	number &  	maximum & maximum & size & size & size  \\ 
%		& of   & memory & number of & of & of & of 
%%		& frames & duration & distractors & training set & validation set & test set \\
%		\midrule
%		Canonical & 4 & 3 & 1 & 10000320 & 500016 & 500016 \\	
%		Hard  & 8 & 7 & 10 & 10000320 & 500016  & 500016 \\
%		\bottomrule	
%	\end{tabular}
%	\caption{Details of the Canonical and Hard variants of the COG dataset.}
%	\label{tab:cog_variants_supplement}
%\end{table}

\clearpage
\section{Complete COG results}
\label{sec:cog-all-results}

\begin{table*}[htb]
	\centering
	\begin{adjustbox}{width=\textwidth}
		\begin{tabular}{l c c c c c c c c c}
			\toprule
			Model & \multicolumn{4}{c}{SAMNet} &  &\multicolumn{4}{c}{Baseline Model} \\
			\cmidrule(lr){2-5} 
			\cmidrule(lr){7-10} 
			&&&&& & paper & code & code & paper\\
			\cmidrule(lr){7-10} 
			Trained on       & canonical & canonical & canonical & hard &           &  canonical  & canonical  & canonical & hard \\ 
			Fine tuned on  & - & - & hard  & - &           & -   & - & hard & - \\ 
			Tested on        & canonical & hard & hard & hard &            &canonical  & hard & hard & hard  \\ 
			\midrule			
			Overall accuracy & 98.0 & 91.6 & 96.5  & 96.1 &           & 97.6  & 65.9 & 78.1& 80.1 \\ 
			\midrule 
			AndCompareColor			&	93.5	&	82.7	&	89.2	&	80.6	&	&	81.9	&	57.1	&	60.7	&	51.4	 \\
			AndCompareShape			&	93.2	&	83.7	&	89.7	&	80.1	&	&	80.0	&	53.1	&	50.3	&	50.7 \\
			AndSimpleCompareColor		&	99.2	&	85.3	&	97.6	&	99.4	&	&	99.7	&	53.4	&	77.1	&	78.2 \\
			AndSimpleCompareShape		&	99.2	&	85.8	&	97.6	&	99.2	&	&	100.0	&	56.7	&	79.3	&	77.9 \\
			CompareColor			&	98.1	&	89.3	&	95.9	&	99.7	&	&	99.2	&	56.1	&	67.9	&	50.1 \\
			CompareShape			&	98.0	&	89.7	&	95.9	&	99.2	&	&	99.4	&	66.8	&	65.4	&	50.5	 \\
			Exist				&	100.0	&	99.7	&	99.8	&	99.8	&	&	100.0	&	63.5	&	96.1	&	99.3 \\
			ExistColor			&	100.0	&	99.6	&	99.9	&	99.9	&	&	99.0	&	70.9	&	99	&	89.8 \\
			ExistColorOf			&	99.9	&	95.5	&	99.7	&	99.8	&	&	99.7	&	51.5	&	76.1	&	73.1 \\
			ExistColorSpace			&	94.1	&	88.8	&	91.0	&	90.8	&	&	98.9	&	72.8	&	77.3	&	89.2 \\
			ExistLastColorSameShape		&	99.5	&	99.4	&	99.4	&	98.0	&	&	100.0	&	65.0	&	62.5	&	50.4 \\
			ExistLastObjectSameObject	&	97.3	&	97.5	&	97.7	&	97.5	&	&	98.0	&	77.5	&	61.7	&	60.2 \\
			ExistLastShapeSameColor		&	98.2	&	98.5	&	98.8	&	97.5	&	&	100.0	&	87.8	&	60.4	&	50.3 \\
			ExistShape			&	100.0	&	99.5	&	100.0	&	100.0	&	&	100.0	&	77.1	&	98.2	&	92.5 \\
			ExistShapeOf			&	99.4	&	95.9	&	99.2	&	99.2	&	&	100.0	&	52.7	&	74.7	&	72.70 \\
			ExistShapeSpace			&	93.4	&	87.5	&	91.1	&	90.5	&	&	97.7	&	70	&	89.8	&	89.80 \\
			ExistSpace			&	95.3	&	89.7	&	93.2	&	93.3	&	&	98.9	&	71.1	&	88.1	&	92.8 \\
			GetColor			&	100.0	&	95.8	&	99.9	&	100.0	&	&	100.0	&	71.4	&	83.1	&	97.9 \\
			GetColorSpace			&	98.0	&	90.0	&	95.0	&	95.4	&	&	98.2	&	71.8	&	73.	&	92.3 \\
			GetShape			&	100.0	&	97.3	&	99.9	&	99.9	&	&	100.0	&	83.5	&	89.2	&	97.1	 \\
			GetShapeSpace			&	97.5	&	89.4	&	93.9	&	94.3	&	&	98.1	&	78.7	&	77.3	&	90.3 \\
			SimpleCompareShape		&	99.9	&	91.4	&	99.7	&	99.9	&	&	100.0	&	67.7	&	96.7	&	99.3 \\
			SimpleCompareColor		&	100.0	&	91.6	&	99.80	&	99.9	&	&	100.0	&	64.2	&	90.4	&	99.3 \\
			\bottomrule
		\end{tabular}
	\end{adjustbox}
	\caption{COG test set accuracies for SAMNet and the baseline model~\cite{yang2018dataset}. The `paper' columns for the baseline model denote results taken from~\cite{yang2018dataset} along with some further author clarifications (private communication) to obtain the performance on individual task types while `code' denotes results of our experiments using their implementation~\cite{yang2018implement}.}
	\label{tab:all-results}
\end{table*}

\clearpage
\section{Complete CLEVR-CoGenT results}
\label{sec:full-cogent-results}

\begin{table*}[htb]
	\centering
	\begin{adjustbox}{width=\columnwidth}
	\begin{tabular}{c c c c c c c }\toprule
		\multicolumn{1}{c }{\textbf{}} & \multicolumn{6}{c }{\textbf{Test Accuracy (\%)}} \\  
		\multicolumn{1}{c }{\textbf{}} & \multicolumn{6}{c }{on \texttt{valA} -- 150,000 samples} \\ 
		\cmidrule(lr){2-7}
		\multicolumn{1}{ c }{\textbf{Experiments}} & \textbf{Overall}& \textbf{Exist}  & \textbf{Count} & \textbf{CompareInteger} & \textbf{CompareAttribute} & \textbf{QueryAttribute}\\ 
		\cmidrule(lr){1-1}
		\cmidrule(lr){2-7}
		\multicolumn{1}{ c }{Exist only} & 26.07 & \emph{74.20}	& 0.0	& 59.79	& 59.15 & 0.0 \\ 
		\multicolumn{1}{ c }{Count only} & 14.89  & 0.0	& \emph{62.78}	& 0.0 & 0.0 & 0.0 \\ 
		\multicolumn{1}{ c }{CompareInteger only} & 23.44 & 48.15	& 0.0	& \emph{77.98}	& 54.08 & 0.0 \\ 
		\multicolumn{1}{ c }{CompareAttribute only} & 23.50 & 51.93	& 0.0 & 59.06 & \emph{61.73} & 0.0 \\ 
		\multicolumn{1}{ c }{QueryAttribute only} & 34.64 	& 0.0	& 0.0	& 0.0 & 0.0 & \emph{97.08} \\ 		
		\cmidrule(lr){1-1}
		\cmidrule(lr){2-7}

		\multicolumn{1}{ c }{All tasks} & 95.32 & 98.4 	& 86.75	& 96.0	& 98.65	& 98.02 \\ 
		\cmidrule(lr){1-1}
		\cmidrule(lr){2-7}

		\multicolumn{1}{ c }{All tasks but Exist} & 90.33 	& \emph{60.42}	& 86.12	& 96.18	& 98.52 & 98.60 \\ 
		\multicolumn{1}{ c }{All tasks but Count} & 74.59 	& 97.51	& \emph{0.0}	& 94.37	& 98.42 & 98.47 \\ 
		\multicolumn{1}{ c }{All tasks but CompareInteger} & 91.53 	& 98.22	& 86.09	& \emph{56.35}	& 98.78 & 98.40 \\ 
		\multicolumn{1}{ c }{All tasks but CompareAttribute} & 81.92 	& 98.32	& 86.36	& 95.86	& \emph{23.18} & 98.44 \\ 
		\multicolumn{1}{ c }{All tasks but QueryAttribute} & 42.17 	& 76.79	& 54.64	& 79.87 & 64.13 & \emph{0.0} \\

		\cmidrule(lr){1-1}
		\cmidrule(lr){2-7}

		\multicolumn{1}{ c }{\textit{Trained on all tasks -- Finetune on $t$}} & &  & &  & &\\ 
		\multicolumn{1}{ c }{Exist} & 94.16 & \emph{98.04} & 82.96 & 95.02 & 98.12 & 97.97 \\ 
		\multicolumn{1}{ c }{Count} & 95.10 & 96.46 & \emph{88.28} & 94.78 & 97.01 & 98.26 \\
		\multicolumn{1}{ c }{CompareInteger} & 95.31 & 98.39 & 86.56 & \emph{96.07} & 98.70 & 98.09 \\
		\multicolumn{1}{ c }{CompareAttribute} & 95.31 & 98.40 & 86.78 & 96.00 & \emph{98.66} & 98.04 \\
		\multicolumn{1}{ c }{QueryAttribute} & 93.49 & 97.45 & 82.09 & 92.23 & 97.85& \emph{97.76} \\

		\bottomrule
	\end{tabular}
	\end{adjustbox}
	\caption{Complete set of results for SAMNet on CLEVR-CoGenT.}
	\label{tab:CoGenT-results}
\end{table*}