### 2019-11-22 Transfer_Learning/arxiv_version/arxiv_paper_main.tex
Title + author Figure placement

\documentclass[11pt]{article}

\usepackage[letterpaper,margin=1in]{geometry}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsmath, amssymb, amsfonts,bm}
\usepackage{amsthm}

% Include other packages here, before hyperref.
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{xcolor}

\usepackage{hyperref}
\hypersetup{
	colorlinks,
	breaklinks,
	linkcolor={red!50!black},
	citecolor={blue!50!black},
	urlcolor={blue!80!black},
}

\usepackage{cleveref}
\crefname{section}{Section}{Sections}
\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{appendix}{Appendix}{Appendices}
\crefname{definition}{Definition}{Definitions}

\newtheorem*{definition}{Definition}

\usepackage{enumitem}
\setlist{leftmargin=*}

%% User's macros %%
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\cF}{\mathcal{F}}

\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vc}{\vec{c}}
\newcommand{\vso}{\vec{so}}
\newcommand{\vone}{\vec{1}}
\newcommand{\proj}{\mathrm{proj}}
\newcommand{\vx}{\vec{x}}
\newcommand{\vy}{\vec{y}}

\newcommand{\imatch}{g^{\mathrm{v}}}
\newcommand{\mmatch}{g^{\mathrm{m}}}

\newcommand{\doadd}{h^{\mathrm{a}}}
\newcommand{\doreplace}{h^{\mathrm{r}}}

\newcommand{\tlast}{\tau^{\mathrm{last}}}
\newcommand{\tlatest}{\tau^{\mathrm{latest}}}
\newcommand{\tnow}{\tau^{\mathrm{now}}}
\newcommand{\tnone}{\tau^{\mathrm{none}}}

\newcommand{\rhead}{\vec{rh}}
\newcommand{\whead}{\vec{wh}}

\newcommand{\uX}{\underline{X}}

\newcommand{\tk}[1]{\textcolor{red}{TK: #1}}
\newcommand{\ao}[1]{\textcolor{green}{AO: #1}}
\newcommand{\tsj}[1]{\textcolor{magenta}{TSJ: #1}}
\newcommand{\vm}[1]{\textcolor{blue}{VM: #1}}

%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%% TITLE
\title{Transfer Learning in Visual and Relational Reasoning}

%%%%%%%%% AUTHORS
% WILL NEED TO REARRANGE THE AUTHORS LIST & GET CORRECT INSTITUTIONS / EMAILS
\author{%
T.S. Jayram\thanks{\textit{Almaden Research Center, San Jose, CA, USA}.} \hspace{1cm} \and  
Vincent Marois\thanks{\textit{Thomas J. Watson Research Center, Yorktown Heights, NY, USA}.} \hspace{1cm} \and 
Tomasz Kornuta\thanks{Currently at Nvidia. \textit{Santa Clara, CA, USA}.} \and
\hspace{2cm} Vincent Albouy\footnotemark[1]  \hspace{1cm} \and 
Emre Sevgen\thanks{Currently at Citrine Informatics}  \hspace{1cm} \and 
Ahmet S. Ozcan\footnotemark[1] \and \\
IBM Research AI
}

\date{}

\maketitle

%%%%%%%%% ABSTRACT
\begin{abstract}
  Transfer learning is becoming the de facto solution for vision and text encoders in the front-end processing of machine learning solutions. Utilizing vast amounts of knowledge in pre-trained models and subsequent fine-tuning allows achieving better performance in domains where labeled data is limited. In this paper, we analyze the efficiency of transfer learning in visual reasoning by introducing a new model (SAMNet) and testing it on two datasets: COG and CLEVR.  Our new model achieves state-of-the-art accuracy on COG and shows significantly better generalization capabilities compared to the baseline. We also formalize a taxonomy of transfer learning for visual reasoning around three axes: feature, temporal, and reasoning transfer. Based on extensive experimentation of transfer learning on each of the two datasets, we show the performance of the new model along each axis.
\end{abstract}

%%%%%%%%% BODY TEXT
\input{introduction}

\input{related_work}

\input{transfer_learning}

\input{samnet}

\input{experiments}

\input{summary}

\bibliographystyle{alpha}
\bibliography{../bibliography}

\newpage
\appendix

\input{appendix}

\end{document}