### 2020-02-05 Transfer_Learning/cvpr_rebuttal/egrebuttal.tex
Added cvpr_rebuttal with LateX template.

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\def\cvprPaperID{7813}
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\begin{document}

\title{Transfer Learning in Visual and Relational Reasoning -- Rebuttal}

\maketitle
\thispagestyle{empty}

\section{Introduction}

{\small
\bibliographystyle{ieee}
\bibliography{}
}

\end{document}