### 2020-02-05 Transfer_Learning/cvpr_2nd_submission/related_work.tex

\section{Related work}

\tk{That section will focus \textbf{only} on transfer learning}

In Computer Vision, it is now standard practice to pretrain an image encoder (such as VGG~\cite{simonyan2014very} or ResNet~\cite{he2016deep}) on large-scale datasets (such as ImageNet~\cite{deng2009imagenet}), and reuse the weights in unrelated domains and tasks, such as segmentation of cars~\cite{iglovikov2018ternausnet} or Visual Question Answering (VQA) in a medical domain~\cite{kornuta2019leveraging}.
Such performance improvements are appealing, especially in cases where both the domain (natural vs. medical images) and the task (image classification vs. image segmentation vs VQA) change significantly. %This type of transfer learning is qualified as \emph{inductive}, according to the taxonomy from~\cite{pan2009survey}.

Similar developments have emerged in the Natural Language Processing (NLP) community.
Using shallow word embeddings, such as word2vec~\cite{mikolov2013distributed} or GloVe~\cite{pennington2014glove}, pretrained on large corpuses from e.g.\ Wikipedia or Twitter, has become a standard procedure when working with different NLP domains and tasks.
Recently, there is a clear, growing trend of utilization of deep contextualized word representations such as ELMo~\cite{peters2018deep} (based on bidirectional LSTMs~\cite{hochreiter1997long}) or BERT~\cite{devlin2018bert} (based on the Transformer~\cite{vaswani2017attention} architecture), where entire deep networks (not just the input layer) are pretrained on very large corporas.
In analogy to pretrained image encoders, the NLP community has also started to create model repositories, some with dozens of pretrained models ready to be downloaded and used. HuggingFace~\cite{wolf2019transformers} is one of the most notable examples.

The success of transfer learning raises several research questions, such as the characteristics which make a dataset more favorable to be used in pretraining (notably ImageNet~\cite{huh2016makes}), or regarding the observed performance correlation of models with different architectures between the source and target domains~\cite{kornblith2019better}.
One of the most systematic works in this area is the computational taxonomic map for task transfer learning~\cite{zamir2018taskonomy}, which aimed at discovering the dependencies between twenty-six 2D, 2.5D, 3D, and semantic computer vision tasks.

\tk{missing review: TL in visual reasoning, refer to CoGenT and COG}

\tk{after reading that section the reader should end up with a conclusion that: there are no good models for TL in VR and, moreover, the datasets are randomly testing this or that, there is no theoretical framework showing that do they mean/bigger picture is missing}