### 2020-02-05 Transfer_Learning/cvpr_2nd_submission/samnet.tex

\section{Selective Attention Memory (SAM) Network}

\tk{Description of the SAMNet}