### 2020-02-05 Transfer_Learning/cvpr_2nd_submission/transfer_learning.tex

\section{Transfer Learning}

\tk{description of the theoretical framework being the foundation driving our experiments}

\subsection{Feature transfer}

\subsection{Temporal transfer}

\subsection{Reasoning transfer}

\tk{experiments should follow that order AND should be consistent with the used terminology}