### 2020-02-06 Transfer_Learning/cvpr_rebuttal/egrebuttal.tex
Updated rebuttal

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\tk}[1]{\textcolor{red}{TK: #1}}
\newcommand{\ao}[1]{\textcolor{green}{AO: #1}}
\newcommand{\tsj}[1]{\textcolor{magenta}{TSJ: #1}}
\newcommand{\vm}[1]{\textcolor{blue}{VM: #1}}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,colorlinks,bookmarks=false]{hyperref}

\def\cvprPaperID{7813}
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\begin{document}

\title{Transfer Learning in Visual and Relational Reasoning -- Rebuttal}
\maketitle
\thispagestyle{empty}

We would like to thank all three reviewers for their time to read our manuscript and providing us useful feedback.  Reviewers agree on the strengths of the paper, namely the introduction of a new model which out-performed the baselines, and commonly point out to the same weaknesses, which is about the writing, motivation and organization of the paper.   The specific recommendations about how to improve the paper were extremely valuable for us.  Based on these, we overhauled the paper and re-wrote most of it (please see the revised version at the end of this document).    Below, we also provide answers to the technical questions raised by the reviewers.

Here are the changes we have made:
\begin{enumerate}
	\item Re-wrote the \textbf{Abstract} to clarify the problem statement and our contribution.
	\item Re-wrote most of the \textbf{Introduction}, shortened it and sharpened the focus about the main contributions.
	\item \textbf{Related Work} section focused more on Transfer learning, and the deficiencies of existing solutions.
	\item We moved the \textbf{Transfer Learning} section just before the Experiments to tie the theoretical framework to the experiments.
	\item \textbf{Model}:  We added a \textbf{new figure} to show the details of SAMNet and depicted the variables to connect to the equations.  We also improved the description of the Model based on Rev.2's recommendations.
	\item We \textbf{added baselines} for CLEVR/CoGenT in Figure X.
	\item We rearranged the order of \textbf{Experiments} to better align with the theoretical framework.
	\item In \textbf{Supplementary} section, we added snapshots from our visualization tool to show the behavior of the model in terms of visual attention, the retrieval of objects from the frame and the use of memory.
\end{enumerate}

We appreciate the time of the reviewers for a second time and hope that they would change their rating to give this work a chance to reach the broader CVPR audience.

\section{Reviewer \#1 questions}

We agree that further ablation studies would be valuable however, per CVPR guidelines, we are not able to preform new experiments during rebuttal.

We pledge to make our code available prior to the conference, however we do not have necessary the approvals yet before the rebuttal period ends.

We've added baselines to the feature transfer. Note that some of the transfer learning experiment settings are novel with no external baselines. We hope that future studies will use our methodology and compare against our model.

\section{Reviewer \#2 questions}

\begin{enumerate}
  \item What are $vo_tt$ and $mo_t$ and how they differ (lines 400-403)? What is the difference between both objects? \vm{$vo_t$ is the result of a retrieval operation on the content of the current input frame. $mo_t$ is the result of a similar operation on the memory content. The difference between both objects is thus their source.}

  \item How the temporal classifier $\tau_t$ is trained and used? Do you use its logits or classes directly (lines 389-397)?
  \item What are $va_t$ and in general all other symbols. How they are computed. \vm{$va_t$ is the visual attention vector, as defined on lines 412 - 413. $va_t$ is the result of a softmax layer}.
  \item What is pseudo-attention (line 453)? Why such a name? \vm{$w_t$ is called a pseudo-attention vector as its norm may not always be 1 (just like a probability vector), but rather vary between 0 (no change) to 1 (complete change)}.
  \item What are reasoning operations (lines 389-391)? \vm{One iteration of the SAM Cell is considered to be a reasoning operation, as defined on lines 376 - 377}.
  \item How many 'reasoning operations' (what is k)? \vm{8 reasoning steps, as noted on lines 532-533.}
  \item Figure 4 shows results on CLEVR-CoGenT. CLEVR-CoGenT uses a transfer between objects of type A and objects of type B (e.g. different combination of shapes and colors). Is this used here? Or this is standard CLEVR results? What are the results on the regular CLEVR, is it 95\% from the supp. material? \vm{The results presented in Fig 4 are reported on the condition A of CLEVR-CoGenT. No results are reported for CLEVR.}
  \item Why the paper doesn't compare to other methods on CLEVR?
\end{enumerate}

% References
% {\small
% \bibliographystyle{ieee}
% \bibliography{}
% }

\end{document}