### 2020-03-03 Transfer_Learning/eccv_submission/eccv_supplementary_material.tex

\documentclass[runningheads]{llncs}

\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{xcolor}
\usepackage{comment}

% INITIAL SUBMISSION - The following two lines are NOT commented
% CAMERA READY - Comment OUT the following two lines
\usepackage{ruler}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

\usepackage{hyperref}
\hypersetup{
	colorlinks,
	breaklinks,
	linkcolor={red!50!black},
	citecolor={blue!50!black},
	urlcolor={blue!80!black},
}

\usepackage{cleveref}
\crefname{table}{Table}{Tables}

\begin{document}

%%%%%%%%% TITLE
\pagestyle{headings}
\mainmatter
\def\ECCVSubNumber{7169}  % Insert your submission number here

\title{Transfer Learning in Visual and Relational Reasoning -- Supplementary Material}

% INITIAL SUBMISSION
%\begin{comment}
\titlerunning{ECCV-20 submission ID \ECCVSubNumber}
\authorrunning{ECCV-20 submission ID \ECCVSubNumber}
\author{Anonymous ECCV submission}
\institute{Paper ID \ECCVSubNumber}
%\end{comment}

% CAMERA READY SUBMISSION
\begin{comment}
\author{T.S. Jayram\inst{1} \and  Vincent Marois\inst{2} \and Vincent Albouy\thanks{Work done while at IBM Research AI, Almaden Research Center} \and Tomasz Kornuta\inst{3} \and Ahmet Ozcan\inst{1}}

\authorrunning{T.S. Jayram et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

\institute{IBM Research AI, Almaden Research Center, San Jose, CA, USA\\ \email{\{jayram,  asozcan\}@us.ibm.com}
      \and IBM Research AI, Thomas J. Watson Research Center, York Town, NY, USA\\ \email{vincent.marois@ibm.com}
      \and NVIDIA, Santa Clara, CA, USA\\ \email{tkornuta@nvidia.com}}
\end{comment}

\maketitle

\input{appendix}

\bigskip

\clearpage
\bibliographystyle{splncs04}
\bibliography{../bibliography}

\end{document}